---
title: "Subjects and Demographics"
author: "Luisa Balzus"
date: "26 Juli 2019"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  tidy = TRUE,
  message = FALSE,
  warning = FALSE,
  dpi = 150
)
```


```{r libraries, include = FALSE}
library(pastecs)    # for descriptive stats
library(foreign)    # for reading in SPSS file
library(knitr)      # for nice tables in html
library(kableExtra) # for nice tables in html

# clear environment
rm(list=ls())

# force R to not use exponential notation
options(scipen = 999)

# reading in datafile
subject_data <- read.spss("P:/Luisa_Balzus/1_PhD_Project/6_ModERN_Behavioral_Study/participant_data.sav", to.data.frame = TRUE)
```
<br>
<br>

# Exclusion of Subjects 

* 31 subjects were assessed. 

* 1 subject was excluded due to a current mental illness and an insufficient number of errors. 

* One further subject (subject 3) had only 2 valid responses to positive words after an error (after an error, this subject classified only 33% of positive words correctly - in comparision to 100% of negative words). However, the criterion used by Aarts (minimum 10 errors) is fulfilled, so the subject does not have to be excluded.

* It remains to be seen, whether subjects have to be excluded as SCR non-responders. 

<br>
<br>
<br>

# Exclusion of Trials 

* 1: invalid rt in gng and word excluded in all analyses to have consistency

* 2: for priming: additionally to 1 excluded wrong keys / misses in gng or words; incorrect responses to words (for RT)

* 3: for scr / priming-scr: additionally to 2 excluded CI, trials followed or preceded by FA / wrong key

* 4: for gng: additionally to 1 excluded wrong keys / misses / CI in gng


<br>
<br>
<br>


# Overview Subjects 

<br>

* N = 30 (22 female, age: M = 22.03, SD = 5.46)

* native German speakers

* all right handed, normal or corrected-to-normal vision

* HU psychology students, participated for course credit

* 7 with past mental issues 

    * I compared ANOVAs for response time and accuracy as well as post hoc tests between the sample including and excluding these subjects. The results do not differ at
    all.
    
    * I also ran *t* tests comparing priming effect in response time and accuracy after FA and FH between the 7 subjects with past mental issues and the 23 subjects
    without. They do not differ in priming effect after FA in response time and accuracy and priming effect after FH in response time and accuracy.



<br>
```{r demographic}

kable(subject_data[,c("CODE","Alter","Geschlecht","Muttersprache","Haendigkeit","Anzahl_verbales_Speeden","Psych_Behandlung","Psych_Behandlung_Grund")], caption = "Subject Data")  %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), font_size = 10, full_width = T, position = "left")

age <- round(stat.desc(subject_data$Alter,basic=F),digits = 2)
````